### 2020-01-19 README.Rmd
still trying to get binder working + docker

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Dendrochronological dates confirm a Late Prehistoric population decline in the American Southwest derived from radiocarbon dates"
```

# robinson2020
[![Travis-CI Build Status](https://travis-ci.org/bocinsky/robinson2020.svg?branch=master)](https://travis-ci.org/bocinsky/robinson2020)
[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/bocinsky/robinson2020/master?urlpath=rstudio)

This repository contains the data and code for our paper:

> Robinson, Erick [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0002-0789-3724), R. Kyle Bocinsky [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0003-1862-3428), Jacob Freeman, Darcy Bird [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0003-3466-6284), Robert L. Kelly, (2020). _`r Title`_. Under review.

<!-- Our pre-print is online here: -->

<!-- > Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx> -->

### How to cite

Please cite this compendium as:

> Bocinsky, R. Kyle and Erick Robinson, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`.

## Contents

The **analysis** directory contains:

- [:file_folder: paper](/analysis/paper): R Markdown source document for the manuscript. Includes code to reproduce the figures and tables generated by the analysis. It also has a rendered version, `paper.html`, suitable for reading (the code is replaced by figures and tables in this file).
- [:file_folder: data](/analysis/data): Data used in the analysis.  
- [:file_folder: figures](/analysis/figures): Plots and other illustrations

## How to run in your broswer or download and run locally

The simplest way to explore the text, code and data is to click on [binder](https://mybinder.org/v2/gh/bocinsky/robinson2020/master?urlpath=rstudio) to open an instance of RStudio in your browser, which will have the compendium files ready to work with. Binder uses [rocker-project.org Docker] images to ensure a consistent and reproducible computational environment. These Docker images can also be used locally.

You can download the compendium as a zip from from this URL: [master.zip](/archive/master.zip). After unzipping, open the `Proj` file in RStudio, and run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the [DESCRIPTION](/DESCRIPTION) file). Then open `paper.Rmd` and knit to produce the `paper.docx`.

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** [GNU GPLv3](LICENSE.md)

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### Acknowledgements

This compendium was created using the [`rrtools` package by Ben Marwick](https://github.com/benmarwick/rrtools), which is ✨pure magic ✨for doing reproducible research.